---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# imcover

<!-- badges: start -->
<!-- badges: end -->

## Testing package of immunisation coverage modelling in R.

The `imcover` package is not yet on CRAN and can be installed from Github. 

Because this is still in a private repository, you must first create an
authentication token. To do this, log in to github.com and got to the Developer
settings:
[https://github.com/setting/tokens](https://github.com/setting/tokens).

1. Select the button to "Generate new token" at the top of the page.
2. Optionally type a note to remind yourself what this token is.
3. Set an expiration date
4. Check the box for **repo** scope to give access to private repositories
5. Copy your token somewhere safe and don't share it! You will need it for the
next step

Now, in `R` you can install the package. In the following command insert your
token as a string:
```{r, install, eval = FALSE}
devtools::install_github('wpgp/imcover@main', auth_token = 'YOUR TOKEN')
```

The build process takes some time because it is compiling the C++ code for the
`Stan` models. It may also ask you to install some additional dependencies.

