---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# imcover: Immunisation coverage modelling in R

<!-- badges: start -->
<!-- badges: end -->

The goal of the `imcover` package is to provide access to data download,
processing and modelling tools to support a Bayesian statistical modelling
approach to generate national estimates of immunization coverage from multiple
time series of data on coverage. `imcover` is built as part of the open-source
statistical computing and modelling language, `R`
([https://cran.r-project.org/](https://cran.r-project.org/)). This package is
designed to support broadly replicable and reproducible analyses of immunization
coverage.

## Statistical modelling software

The core of `imcover` is the functionality to fit a Bayesian statistical model
of multiple time series. The sources of coverage data (in this example
administrative, official and surveys) are taken as multiple, partial estimates
of the true, unobserved immunization coverage in a country. A Bayesian
estimation approach allows us to incorporate these multiple datasets, place
prior beliefs on which sources are more reliable, share information between
countries, and to quantify uncertainty in our estimate of the latent
immunization coverage.

`imcover` provides an interface to [`Stan`
(https://mc-stan.org/)](https://mc-stan.org/) for statistical computation. This
means that, in addition to `imcover`, many of the tools for assessing model
performance and visualizing results from `Stan` will work for `imcover` results.
However, users must have `Stan` installed and linked with `R` in order to use
`imcover`

To install `Stan`, please follow the instructions for your operating system
described here:
[https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started](https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started)

## Installing `imcover`

The `imcover` package is not yet on CRAN and can be installed from Github. 

Because this is still in a private repository, you must first create an
authentication token. To do this, log in to github.com and got to the Developer
settings:
[https://github.com/setting/tokens](https://github.com/setting/tokens).

1. Select the button to "Generate new token" at the top of the page.
2. Optionally type a note to remind yourself what this token is.
3. Set an expiration date.
4. Check the box for **repo** scope to give access to private repositories
5. Copy your token somewhere safe and don't share it! You will need it for the
next step.

Now, in `R`, you can install the package. In the following command insert your
token as a string:
```{r, install, eval = FALSE}
devtools::install_github('wpgp/imcover@main', auth_token = 'YOUR TOKEN')
```

The build process takes some time because it is compiling the C++ code for the
`Stan` models. It may also ask you to install some additional dependencies.

## Contributions

Feedback and contributions are welcome. Please raise or respond to an issue, or
create a new branch to develop a feature/modification and submit a pull request.


