---
title: "Model-based approaches to immunization coverage estimations in `R`: An introduction to the `imcover` package"
author: Chris Jochem (w.c.jochem@soton.ac.uk)
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
vignette: >
  %\VignetteIndexEntry{Introduction to `imcover`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy = FALSE
)
```

National estimates of immunization coverage are crucial for monitoring and
evaluating coverage levels and trends, as well as immunization goals and
targets, at the national and international levels. There are multiple potential
sources of information related to immunization coverage, and the challenge is to
develop a statistically-principled method to make use of different data sources
while accounting for their uncertainties.

The goal of the `imcover` package is to provide access to data download,
processing and modelling tools and to support a Bayesian statistical modelling
approach to generate national estimates of immunization coverage from multiple
time series of data on coverage. `imcover` is built as part of the open-source
statistical computing and modelling language, `R`
([https://cran.r-project.org/](https://cran.r-project.org/)). This package is
designed to support broadly replicable and reproducible analyses of immunization
coverage.

This vignette will demonstrate some of the core functionality in the package,
including:

* Downloading input data sources on coverage
* Cleaning/processing/harmonizing data sources
* Fitting a Bayesian statistical model and model outputs
* Generating national coverage estimates and visualizing results

### Installation and set-up

`imcover` is not (yet) available for download from CRAN. It is currently in a
private GitHub repository during development. Access can be granted to users for
"beta testing." Once downloaded, a user accesses the code library through `R`.

```{r setup}
# load the package within the R environment
library(imcover)
```

All functions are documented and the help and examples can be accessed within
`R` using the standard command `help` or `?`.

`imcover` was developed by [WorldPop at the University of
Southampton (https://www.worldpop.org/)](https://www.worldpop.org/).

## Accessing and downloading data

If a user does not have a dataset on their local computer, the first step in
analysis is to acquire standardised data. Within `imcover` there is
functionality to download several datasets from the WHO Immunization data portal
([https://immunizationdata.who.int/listing.html](https://immunizationdata.who.int/listing.html)).
These datasets include official coverage reports, administrative records, as
well as household survey datasets.

```{r dataDownload}
# download administrative and official records
cov <- download_coverage()

# download survey records
svy <- download_survey()
```

As part of the download process, by default, several pre-processing steps are
done to, e.g. reconcile differences in column names. In the case of survey
datasets, a recall-bias adjustment is applied. The default processing behaviour
can be changed by modifying the argument supplied to the download functions.
Additionally, all the processing steps can be carried out using functions
available in `imcover` if advanced users want more control over pre-processing.

The data files available from the WHO website require some additional cleaning
before analysis. Notice that the data objects created by `imcover` can work with
all standard `R` commands for selecting and modifying records.

```{r dataCleaning}
# Further data cleaning of immunization records
# drop some record categories (PAB, HPV, WUENIC)
cov <- cov[cov$coverage_category %in% c("ADMIN", "OFFICIAL"), ]
cov$coverage_category <- tolower(cov$coverage_category)  # clean-up

# create a combined dataset
dat <- rbind(cov, svy)

# remove records with missing coverage values
dat <- dat[!is.na(dat$coverage), ]

# mismatch in vaccine names between coverage and survey datasets
dat[dat$antigen == 'DTPCV1', 'antigen'] <- 'DTP1'
dat[dat$antigen == 'DTPCV2', 'antigen'] <- 'DTP2'
dat[dat$antigen == 'DTPCV3', 'antigen'] <- 'DTP3'
```

## `ic` data objects

A unique feature of `imcover` is that it creates and works with a custom data
extension, which we refer to as an immunization coverage data frame or an
`ic.df`. This type of object has all the functionality of a standard `R` data
frame while having additional functionality to store attributes to make it
easier to work with the immunization coverage data and to keep track of what
processing has been done to the data.

```{r inspectObj}
# inspect a random sample of records
print(dat[sample(1:nrow(dat), size = 6), ])
```

```{r icObj}
# note the type object
class(dat)
```

For example, the `ic.df` objects have helper functions to tabulate records,

```{r listVax}
# show all records
list_vaccines(dat)  # note the download files contain many vaccine
```

Or to easily filter the records. Here we will subset the data for five core
vaccines and a subset of time points.

```{r filterObj}
# subset the records by vaccine and year for all countries
dat <- ic_filter(dat, 
                 vaccine = c("DTP1", "DTP3", "MCV1", "MCV2", "PCV3"),
                 time = 2000:2018)

# cross-tabulations of source and vaccine
table(dat$coverage_category, dat$antigen)
```

Other specialised methods are available to filter out records for a country
after the year of national introduction.

```{r yovi}
# subset records based on year of vaccine introduction
dat <- filter_yovi(dat)
  table(dat$year, dat$antigen)  # notice year of introduction for PCV3
```

Some further pre-processing of immunization data has been found to be necessary
in preparation for the statistical modelling. We introduce functions to make it
easy to carry out these steps.

Firstly, some records observe inconsistencies in the levels of coverage between
multi-dose vaccines. To maintain consistency, where coverage of later doses
cannot exceed earlier doses, we model the ratio between first and third dose. In
this example, we only adjust DTP1 and DTP3, but other multi-dose vaccines could
be processed in a similar manner.

```{r adj1}
# adjustment - use ratio for DTP3
dat <- ic_ratio(dat, numerator = 'DTP3', denominator = 'DTP1')
```

The `ic.df` object will now store a note that this processing step has been
carried out so that the ratio is back-transformed and that coverage estimates
and predictions are adjusted appropriately.

Secondly, we need to force coverage estimates to lie between 0% and 100%.

```{r adj2}
# maintain coverage between 0-100%
dat <- ic_adjust(dat, coverage_adj = TRUE)
```

## Model fitting

The core of `imcover` is the functionality to fit a Bayesian statistical model
of multiple time series. The sources of coverage data (in this example
administrative, official and surveys) are taken as multiple, partial estimates
of the true, unobserved immunization coverage in a country. A Bayesian
estimation approach allows us to incorporate these multiple datasets, place
prior beliefs on which sources are more reliable, share information between
countries, and to quantify uncertainty in our estimate of the latent
immunization coverage.

`imcover` provides an interface to [`Stan`
(https://mc-stan.org/)](https://mc-stan.org/) for statistical computation. This
means that, in addition to `imcover`, many of the tools for assessing model
performance and visualizing results from `Stan` will work for `imcover` results.

In this demonstration will start by fitting a model to a single region of data
using `ic_fit` by passing in the `ic.df` object and several arguments that
control the posterior sampling.

```{r fit0, message=FALSE, cache=FALSE, include=FALSE}
options(mc.cores = parallel::detectCores())  # enable parallel processing

# Fit model to a single region of data
fit1 <- ic_fit(dat[dat$region == 'AFRO', ], 
               chains = 2, 
               iter = 200, 
               warmup = 100)
```

```{r fit, message = FALSE, eval = FALSE}
options(mc.cores = parallel::detectCores())  # enable parallel processing

# Fit model to a single region of data
fit1 <- ic_fit(dat[dat$region == 'AFRO', ], 
               chains = 4, 
               iter = 2000, 
               warmup = 1000)
```

### Model outputs

After fitting the model, we can work with the parameter estimates and estimates
of coverage. For example, we can easily make graphs of coverage estimates.
Notice how `imcover` has a generic function defined to specially handle objects
returned by `ic_fit`.

```{r plotCov, fig.width = 5, fig.height = 5}
# Plot of coverage
plot(ic_filter(fit1, vaccine = 'DTP1',
               country = c('BWA', 'SLE', 'GHA', 'NGA')),
     ncol = 2)
```
```{r plotCov2, fig.width = 5, fig.height = 5, fig.cap = "Example results of modelled immunization coverage estimates for selected countries and vaccines."}
plot(ic_filter(fit1, vaccine = 'DTP3',
               country = c('BWA', 'SLE', 'GHA', 'NGA')),
     ncol = 2)
```

The main output of the model is a table of coverage estimates for each country,
vaccine, and time point along with uncertainty around each estimate. We can
extract these data as a table, save them to a data frame, or write them out to a
file for use in a report.

```{r icCover}
# Extracting coverage estimates
ic <- ic_coverage(fit1, 
                  stat = "quantile", # customise the summary function
                  probs = c(0.025, 0.5, 0.975))

head(ic, 10)
```

### Predictions

Once we have the fitted model, we can also use it to predict coverage to time
points in the near future.

```{r plotPred, fig.width = 5, fig.height = 5}
# Predict for future time points (3 years post 2018)
fit1 <- predict(fit1, t = 3)  # update icfit to include 'prediction' info

# Update the graph 
# Automatically include fitted estimates, observed data, and predictions
plot(ic_filter(fit1, vaccine = 'DTP1',
               country = c('BWA', 'SLE', 'GHA', 'NGA')),
     ncol = 2)
```

```{r plotPred2, fig.width = 5, fig.height = 5, fig.cap = "Example reults of estimated immunization coverage with future predictions to the right of the dashed lines. Overlaid with observed data points."}
plot(ic_filter(fit1, vaccine = 'DTP3',
               country = c('BWA', 'SLE', 'GHA', 'NGA')),
     ncol = 2)
```

## Summary

This vignette has introduced the core functionality of the `imcover` package and
its use in modelling national time series of immunization coverage. Providing a
coordinated toolset to download, process, and model immunization data should
allow for consistent analyses whose methods are transparent and replicable.

There are a few additional options in the package. For instance, multiple
regions of data can be supplied to `ic_fit` and then region-specific models can
be executed. In the posterior analyses, population-weighted regional summaries
of coverage can also be calculated. Similar to `ic_coverage`, users can
customize the summary functions. Finally, the `Stan` models allow for
significant flexibility in defining statistical models. The package specifies
reasonable default values for the models to replicate our analyses, but users
can explore the sensitivity of these settings by easily adjusting the parameters
within `R` and re-running the models without having to interact with the `Stan`
code.

Feedback on the package is welcome. Please email Chris Jochem
(w.c.jochem@soton.ac.uk)

---
```{r ending, echo = FALSE}
sessionInfo()
```

